Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-04-19 paper/ijcai16.tex
paper: finalize as much as I can.

\paragraph{MILP Formulation.} This initial formulation is problematic
to solve, as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over
mixed continuous integer variables.  \andrea{However, the problem can
  be reformulated as a mixed integer linear program (MILP) by a
  suitable variable transformation. This technique is rather common in
  operational research, see e.g.~\cite{boutilier2006}.}

\andrea{Our goal is to replace Eq.~(\ref{eq:wxconstr}) with a set of linear
constraints}. In order to do so, we introduce a set of fresh variables
$p^{i,j}_z$ for every $i,j\in[k]$ and $z\in[m]$. Assuming for the time
being that the new variables do satisfy the equation
$p^{i,j}_z = w^i_z x^j_z$, we rewrite the fourth component of the
objective function in terms of the new variables as:
%
$$ \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z$$
%
and, similarly, Eq.~(\ref{eq:wxconstr}) as:
%
\[ \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge \mu \label{eq:pxconstr} \]
%
The fact that $p^{i,j}_z = w^{i}_z x^{j}_z$ is achieved by
setting the following additional constraints. We distinguish between two cases:
(i) $p^{i,i}_z$ and (ii) $p^{i,j}_z$ for $i \ne j$.  Recall that we are
maximizing the margin $\mu$. Now, due to Eq.~(\ref{eq:pxconstr}), the optimizer will
try to keep $p^{i,i}_z$ as large as possible and $p^{i,j}_z$ as small as
possible.

@@

(Case ii) We add an explicit lower bound:
%
$$ p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \} $$
%
If $x^j_z = 1$ the lower bound simplifies to
$\max \{ 0, w^{i}_z \} = w^{i}_z$, due to the non-negativity of
$w^i_z$. Otherwise, if $x^j_z = 0$ then the lower bound becomes
$\max \{ 0, w^{i}_z - w_\text{max} \}$, where the second term is at
most $0$. Since $p^{i,j}_z$ is being minimized, in both cases it will
attain the lower bound, and thus satisfy
$p^{i,j}_z = w^i_z x^j_z$.\footnote{Since $\mu$ is upper-bounded by
  Eq.~(\ref{eq:wyconstr}), in some cases the $p^{i,j}_z$ variables do
  not attain the lower bound. As a consequence, the MILP reformulation
  of Eq.~(\ref{eq:wxconstr}) is a (tight) approximation of the
  original one. This \andrea{can only happen in the case of $k=1$} and
  has no impact on the quality of the solutions.}

@@

We implemented the {\sc SetMargin} algorithm using Python, leveraging Gurobi
6.5.0 for solving the core MILP problem. Both the {\sc SetMargin} source code
and the full experimental setup are available at 
\url{https://db.tt/oqY62odW} \stefano{update the link!}.

@@

In all experiments {\sc SetMargin} uses an internal 5-fold cross-validation procedure to
update the hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5
iterations. The hyperparameters are chosen as to minimize the ranking loss over
the user responses collected so far. $\alpha$ is taken in $\{20, 10, 5, 1\}$,
while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.\footnote{Note
that \stefano{for $k=1$ Eq.~\ref{eq:volatile1} and Eq.~\ref{eq:volatile2} disappear,
so} $\alpha$ can not be taken to be less than $1$, as in this case, the
objective can be increased arbitrarily while keeping the right-hand side of
Eq.~(\ref{eq:wyconstr}) constant, rendering the problem unbounded.}

@@

We presented a max-margin approach for efficient preference
elicitation in large configuration spaces.\footnote{Note that max-margin learning has been proposed before \cite{gajos2005} for preference elicitation, 
but with  rudimental methods for query selection.}
Our approach relies on an
extension of max-margin learning to sets, and is effective in the
generation of a diverse set of configurations that can be used to ask
informative preference queries.  The main advantages of this
elicitation method are 1) ability to provide recommendations in large
configuration problems 2) robustness with respect to erroneous
feedback and 3) ability to encourage sparse utility functions.
Experimental comparisons against state-of-the-art Bayesian preference
elicitation strategies confirm these advantages. Future work includes
extending the approach to truly hybrid scenarios (where real valued
attributes do not depend on categorical ones) and studying its
applicability to other problems, \paolo{(as the identification of Choquet models \cite{Pine2013}}.

>>>>

\paragraph{MILP Formulation.} This initial formulation is problematic
to solve, as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over mixed
continuous integer variables. However, the problem can be reformulated as a
mixed integer linear program (MILP) by a suitable transformation. This
technique is rather common in operational research, see
e.g.~\cite{boutilier2006}.

Our goal is to replace Eq.~(\ref{eq:wxconstr}) with a set of linear
constraints. In order to do so, we introduce a set of fresh variables
$p^{i,j}_z$ for every $i,j\in[k]$ and $z\in[m]$. Assuming for the time
being that the new variables do satisfy the equation
$p^{i,j}_z = w^i_z x^j_z$, we rewrite the fourth component of the
objective function in terms of the new variables as:
%
$$ \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z$$
%
and, similarly, Eq.~(\ref{eq:wxconstr}) as:
%
\[ \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge \mu \label{eq:pxconstr} \]
%
The fact that $p^{i,j}_z = w^{i}_z x^{j}_z$ is achieved by
setting the following additional constraints. We distinguish between two cases:
(i) $p^{i,i}_z$ and (ii) $p^{i,j}_z$ for $i \ne j$.  Recall that we are
maximizing the margin $\mu$. Now, due to Eq.~(\ref{eq:pxconstr}), the optimizer will
try to keep $p^{i,i}_z$ as large as possible and $p^{i,j}_z$ as small as
possible.

@@

(Case ii) We add an explicit lower bound:
%
$$ p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \} $$
%
If $x^j_z = 1$ the lower bound simplifies to
$\max \{ 0, w^{i}_z \} = w^{i}_z$, due to the non-negativity of
$w^i_z$. Otherwise, if $x^j_z = 0$ then the lower bound becomes
$\max \{ 0, w^{i}_z - w_\text{max} \}$, where the second term is at
most $0$. Since $p^{i,j}_z$ is being minimized, in both cases it will
attain the lower bound, and thus satisfy
$p^{i,j}_z = w^i_z x^j_z$.\footnote{Since $\mu$ is upper-bounded by
Eq.~(\ref{eq:wyconstr}), in some cases the $p^{i,j}_z$ variables do not attain
the lower bound. As a consequence, the MILP reformulation of
Eq.~(\ref{eq:wxconstr}) is a (tight) approximation of the original one. This
has no impact on the quality of the solutions.}

@@

We implemented the {\sc SetMargin} algorithm using Python, leveraging Gurobi
6.5.0 for solving the core MILP problem. Both the {\sc SetMargin} source code
and the full experimental setup are available at 
\url{https://github.com/stefanoteso/setmargin}.

@@

In all experiments {\sc SetMargin} uses an internal 5-fold cross-validation procedure to
update the hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5
iterations. The hyperparameters are chosen as to minimize the ranking loss over
the user responses collected so far. $\alpha$ is taken in $\{20, 10, 5, 1\}$,
while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.\footnote{Note
that for $k=1$ Eq.~\ref{eq:volatile1} and Eq.~\ref{eq:volatile2} disappear,
so $\alpha$ can not be taken to be less than $1$, as in this case, the
objective can be increased arbitrarily while keeping the right-hand side of
Eq.~(\ref{eq:wyconstr}) constant, rendering the problem unbounded.}

@@

We presented a max-margin approach for efficient preference
elicitation in large configuration spaces.\footnote{Note that max-margin learning has been proposed before \cite{gajos2005} for preference elicitation, 
but with  rudimental methods for query selection.}
Our approach relies on an
extension of max-margin learning to sets, and is effective in the
generation of a diverse set of configurations that can be used to ask
informative preference queries.  The main advantages of this
elicitation method are 1) ability to provide recommendations in large
configuration problems 2) robustness with respect to erroneous
feedback and 3) ability to encourage sparse utility functions.
Experimental comparisons against state-of-the-art Bayesian preference
elicitation strategies confirm these advantages. Future work includes
extending the approach to truly hybrid scenarios (where real valued
attributes do not depend on categorical ones) and studying its
applicability to other problems, \paolo{as the identification of Choquet models \cite{Pine2013}}.
